---
title: "Week 3 analysis"
author: "Amy"
date: "18/01/2020"
output: html_document
---

We'll begin by loading up the libraries and data we need, as always.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# loading the libraries
library(tidyverse)
library(knitr)
library(here)

# loading the data
loc <- here("gladlysurvey2.csv")
gdata <- read_csv(file=loc)
```

First let's try making a table of people's colours.

```{r sampletablechunk}
colours <- c("grey","brown","navy","white","white","pink","pink")
colourtable <- table(colours)
```

This is kind of ugly. Let's fix it:

```{r sampletablechunk2, results="asis", echo=FALSE}
kable(colourtable, caption="Here is a table of people's colours")
```


Next let's try some cross-tabulation of colours by gender.

```{r samplectchunk, results="asis", echo=FALSE}
genders <- c("female","male","female","male","female","nb","female")
cgtable <- table(genders,colours)
kable(cgtable, caption="Here is a table of people's colours by gender")
```

Finally, let's make a table of some of Gladly's data. First, how many do we have of each species?

```{r specieschunk, results="asis", echo=FALSE}
speciestable <- table(gdata$species)
kable(speciestable, caption="All species in Gladly's data", col.names=c("Species","Frequency"))
```

Let's use pipes and some of our dplyr functions to do some data manipulation. First, we're going to calculate the mean age for each gender using the `group_by()` and `summarise()` functions.

```{r genderagesummary, echo=FALSE}
gdata %>%
  group_by(gender,species) %>%
  summarise(mnAge = mean(age),
            sdAge = sd(age),
            nAge = n()) %>%
  ungroup()
```

Now we're going to do some filtering and arranging

```{r simplefilter, echo=FALSE}
females <- gdata %>%
  filter(gender=="female")

females %>%
  group_by(species) %>%
  summarise(mnAge = mean(age),
            nAge = n()) %>%
  ungroup()
```

Now let's bring it all together.

```{r bringingtogether, echo=FALSE}
gdata %>%
  filter(!is.na(mud)) %>%
  mutate(goodFood = (carrot+cake)/2)  %>%
  group_by(gender) %>%
  summarise(meanGoodFood = mean(goodFood),
            meanBadFood = mean(mud)) %>%
  ungroup()
```

